---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# anrpackageusingc

<!-- badges: start -->
<!-- badges: end -->

The goal of anrpackageusingc is to demonstrate the structure of calling a C function in a package. For a guide on actually writing C code using the R API, see the [appropriate chapter of the R Packages book](https://r-pkgs.org/src.html#clang), the [appropriate chapter of Avanced R](http://adv-r.had.co.nz/C-interface.html) and the [r-internals](https://github.com/hadley/r-internals) collection of unofficial R API documentation.

## Installation

You can install the development version from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/anrpackageusingc")
```

This package is designed so that you can fork, clone, and modify the C code.

## Example

This package contains one function that calls a C function.

```{r example}
library(anrpackageusingc)
my_function("an input object")
```

The R function `my_function()` lives in `R/my-function.R`:

```r
my_function <- function(input) {
  .Call(anrpackageusingc_c_my_function, input)
}
```

The C function `anrpackageusingc_c_my_function()` lives in `src/my-function.c`. I use the prefix `anrpackageusingc_c_...()` when I'm declaring a function that will get registered and called from R because I have a tool that generates `src/init.c` using a regex and because I read something a long time ago that suggested that exported C functions might need unique names (I don't know if this is still true). You can call the functions anything you want but it's helpful if you have a naming convention for all the functions of this type in your package.

```c
SEXP anrpackageusingc_c_my_function(SEXP input_sexp) {
  return R_NilValue;
}
```

It's registered in `src/init.c`. The `1` at the end of the registration line is because the function accepts one argument; the `NULL, NULL, 0` is to signify the end of the array (kind of like how `\0` signifies the end of a null-terminated string).

```c
SEXP anrpackageusingc_c_my_function(SEXP input_sexp);

static const R_CallMethodDef CallEntries[] = {
  {"anrpackageusingc_c_my_function", (DL_FUNC) &anrpackageusingc_c_my_function, 1},
  {NULL, NULL, 0}
};

void R_init_anrpackageusingc(DllInfo *dll) {
  R_registerRoutines(dll, NULL, CallEntries, NULL, NULL);
  R_useDynamicSymbols(dll, FALSE);
}
```

If you want to add a function, you'll have to copy the definition and a line to `CallEntries`. The function can live in any `.c` file of your choosing.

Finally, we need to add the `useDynLib()` line in the `NAMESPACE`. I use the automatically populated roxygen line that happens when you call `usethis::use_package_doc()` and `usethis::use_c("some-c-filename")` for this.

```r
#' @useDynLib anrpackageusingc, .registration = TRUE
```

There are some variations on these components but this should get you started. From scratch, I created this package using a bunch of usethis helpers:

```r
create_package("anrpackageusingc")
# modify Title and Description in DESCRIPTION file
use_package_doc()
use_c("my-function")
# write skeleton of anrpackageusingc_c_my_function() that returns R_NilValue
use_c("init.c")
# copy boilerplate from a previous package using C like this one
use_r("my-function")
# write wrapper using .Call()
use_test("my-function")
# write a test that calls my_function()
use_git()
use_github()
```

## Exercises

1. Modify the body of `anrpackageusingc_c_my_function()` to (1) print something using `Rprintf()` and (2) return an R object. The easiest R objects to create are strings (`Rf_mkString("some text")`), integers (`Rf_ScalarInteger(1234)`), and doubles (`Rf_ScalarReal(1234)`).

2. Add a new C function called `anrpackageusingc_c_my_new_function()` that accepts two parameters (`SEXP item1, SEXP item2`) and returns the first one. Add the function to `init.c` and create an R wrapper called `my_new_function()`.

3. Add a new C function in a new file that adds 2 to a numeric vector (you can use the example from the "Accessing vector data" section of the [appropriate chapter of Avanced R](http://adv-r.had.co.nz/C-interface.html)). Make sure you can call it from R!
